---
title: "p8105_hw2_mgw2144"
output: github_document
date: "2023-09-27"
---

```{r}
library(tidyverse)
library(here)
rm(list=ls())
project.folder=
  paste0(print(here::here()),'/')
```

Problem 1

```{r}
pols_month_df =
  read_csv(paste0(project.folder,"/data/pols-month.csv")) %>% 
  janitor::clean_names()%>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    month = replace(month, month == "01", "jan"),
    month = replace(month, month == "02", "feb"),
    month = replace(month, month == "03", "mar"),
    month = replace(month, month == "04", "apr"),
    month = replace(month, month == "05", "may"),
    month = replace(month, month == "06", "jun"),
    month = replace(month, month == "07", "jul"),
    month = replace(month, month == "08", "aug"),
    month = replace(month, month == "09", "sep"),
    month = replace(month, month == "10", "oct"),
    month = replace(month, month == "11", "nov"),
    month = replace(month, month == "12", "dec")
  ) %>% 
  mutate(president = if_else(prez_gop ==1, "gop", "dem")) %>% 
   select(-day, -prez_gop, -prez_dem)
```

```{r}
snp_df = 
 read_csv(paste0(project.folder,"/data/snp.csv")) %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(
    month = replace(month, month == "1", "jan"),
    month = replace(month, month == "2", "feb"),
    month = replace(month, month == "3", "mar"),
    month = replace(month, month == "4", "apr"),
    month = replace(month, month == "5", "may"),
    month = replace(month, month == "6", "jun"),
    month = replace(month, month == "7", "jul"),
    month = replace(month, month == "8", "aug"),
    month = replace(month, month == "9", "sep"),
    month = replace(month, month == "10", "oct"),
    month = replace(month, month == "11", "nov"),
    month = replace(month, month == "12", "dec")
  ) %>% 
  select(-day) %>% 
  relocate(year, month)
```
```{r}
unemployment_df = 
 read_csv(paste0(project.folder,"/data/unemployment.csv")) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent_unemployed"
  )
```

merge step 

```{r}
pols_snp_merge=
  left_join(pols_month_df, snp_df, by = "year")
```

#### 
will come backt to this
####

Problem 2 
```{r}
mr_trash_wheel=
  readxl::read_excel(paste0(project.folder,"/data/202309 Trash Wheel Collection Data.xlsx"), sheet = 1, range = "A2:N586") %>%
  janitor::clean_names() %>% 
  mutate(
    homes_powered=((weight_tons*500)/1000),
    year = as.integer(year),
    trash_data = "mr_tw")%>% 
  relocate(trash_data)
```

```{r}
professor_trash_wheel=
  readxl::read_excel(paste0(project.folder,"/data/202309 Trash Wheel Collection Data.xlsx"), sheet=2, range = "A2:M108") %>% 
  janitor::clean_names() %>% 
  mutate(homes_powered=(
    (weight_tons*500)/1000),
    year = as.integer(year),
    trash_data = "professor_tw")%>% 
  relocate(trash_data)
```

```{r}
gwynnda_trash_wheel=
  readxl::read_excel(paste0(project.folder,"/data/202309 Trash Wheel Collection Data.xlsx"), sheet=4, range = "A2:L157") %>% 
  janitor::clean_names() %>% 
  mutate(homes_powered=(
    (weight_tons*500)/1000), 
    year = as.integer(year), 
    trash_data = "gwynnda_tw")%>% 
  relocate(trash_data)

gwynnda_july_2021=
  gwynnda_trash_wheel %>% 
  filter(month == "July") %>% 
  filter(year == "2021")
```

```{r}
trash_wheel_total=
  bind_rows(mr_trash_wheel, professor_trash_wheel, gwynnda_trash_wheel) %>%
  janitor::clean_names()
```

The collection of Mr. Trash Wheel datasets records information of a water-wheel vessel tht removed trash from the Inner Harbor in Baltimore, Maryland. The combined dataset created above contains information from the Mr. Trash Wheel, Professor Trash Wheel, and Gywnnda Trash Wheel datasets with a total of `r nrow(trash_wheel_total)` observations. The total weight of trash collected by Professor Trash Wheel was `r sum(pull(professor_trash_wheel, weight_tons))` tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(pull(gwynnda_july_2021, cigarette_butts))`.

Problem 3

```{r}
mci_baseline=
  read_csv(paste0(project.folder,"/data/MCI_baseline.csv"), skip = 1) %>% 
  janitor::clean_names() 

mci_baseline_clean=
  mci_baseline %>% 
  mutate(sex = recode(sex, `1` = "male", `0`="female")) %>% 
  mutate(apoe4 = recode(apoe4, `1`="carrier", `0`="non-carrier")) %>% 
  filter(age_at_onset > current_age | age_at_onset == ".") 

mci_apoe_c=
  mci_baseline_clean %>% 
  filter(sex == "female") %>% 
  filter(apoe4 == "carrier") %>% 
  mutate(apoe4 = recode(apoe4, "carrier" = 1))
  
mci_apoe_nc=
  mci_baseline_clean %>% 
  filter(sex == "female") %>% 
  filter(apoe4 == "non-carrier") %>% 
   mutate(apoe4 = recode(apoe4, "non-carrier" = 1))
```

Important steps in the import process include utilizing the read_csv() function to import the data so that it can be cleaned and manipulated. Important features of the dataset include patient ID number, sex, APOE4 gene carrier status, and age of MCI onset for participants. There were originally 483 participants recruited and `r nrow(mci_baseline_clean)` of these participants were MCI free at baseline. The average baseline age is `r mean(pull(mci_baseline, current_age))` years old. `r (((sum(pull(mci_apoe_c, apoe4)))/((sum(pull(mci_apoe_nc, apoe4)))+(sum(pull(mci_apoe_c, apoe4)))))*100)` percent of women in the study are APOE4 carriers. 

```{r}
mci_amyloid=
  read_csv(paste0(project.folder,"/data/mci_amyloid.csv"), skip = 1) %>% 
  janitor::clean_names() %>% 
  rename("time_0" = baseline) %>%  
  rename("id" = study_id) %>% 
  pivot_longer(
    time_0:time_8,
    names_to = "time",
    values_to = "amyloid_value"
  ) 
```
The import process for this data set involves reading in the csv file for the mci_amyloid data utilizing its relative path name. Some of the prominent features of the dataset include patient ID number, and the time elapsed since the study baseline to the visit where the biomarker Amyloid _ 42/40 ratio was measured marked at baseline and years 2, 4, 6, and 8, and the amyloid value recorded at these times. In cleaning the dataset I changed the column name of participant IDs from "study_id" to "id" to match the "id" column name in the baseline data and changed the "baseline" variable to "time_0". The new dataset has `r nrow(mci_amyloid)` rows and `r ncol(mci_amyloid)` columns. I used the pivot_long function to change the time column names into variables in a new "time" column with their corresponding amyloid values for each unique patient ID number. 

```{r}
mci_unique=
  anti_join(mci_baseline_clean, mci_amyloid, by = "id")
```

There are 8 patients that are only in the baseline or amyloid datasets, patient ids: 14, 49, 92, 179, 268, 304, 389, and 412. These patients were found using the anti_join() function connecting the mci_baseline_clean and mci_amyloid datasets.

```{r}
mci_combine=
  inner_join(mci_baseline_clean, mci_amyloid, by = "id")
```

The dataset "mci_combine" was created by merging only the data on participants in both the "mci_baseline" and "mci_amyloid" datasets throug the inner_join() function. The combined dataset has `r nrow(mci_combine)` rows and `r ncol(mci_combine)` columns. The prominent variables in this dataset are patient id ("id"), age at baseline "current_age", sex, education level ("education"), whether or not they are an APoE4 gene carrier, age at onset of MCI, time in years from baseline when amyliod values were tested "time", and amyloid value at those time intervals. 

###
export the result as a CSV to your data directory. 
###
